The Galactic Midplane Is Not a Plane: Implications for Dynamical Analysis with {\em Gaia} Data and Beyond

### 2019-05-04 paper/ms.tex
Update LSR abbreviatons

In this work we discuss how the imposition of a global axisymmetric coordinate
system based on a local measurement of the midplane location and
Galactocentric distance introduces systematic error in the computation of
actions, especially for stars in the thin disk. Our main point is that the
``local'' Galactic midplane is not the same as the ``global'' Galactic
midplane. For stars relatively far from the Sun, the local midplane location
will not agree with our local midplane extrapolated onto their position, due
to relatively small-scale variations in the density of the gas and stellar
distribution. Converting positions and velocities of more distant stars from a
heliocentric to a Galactocentric coordinate system introduces a systematic
bias in the $z$ coordinate, and hence the actions computed for these stars.
The further from the solar neighborhood the target star, the more likely the
mismatch will result in large systematic uncertainty, especially in the
vertical action $J_z$. A similar
argument applies to any remaining uncertainty in measurements of the
Galactocentric radius of the Sun and to variations in the local standard of
rest (LSR).

@@

\subsection{Local Standard of Rest}
Finally, one must define the local standard of rest (LSR), or mean velocity of
stars near the Sun relative to the Galactic center (which is defined to have
zero velocity), and the velocity of the Sun relative to the LSR. The radial
($U_{\odot}$) and vertical ($W_{\odot}$) components are computed by taking the
mean motions of different stellar groups \citep[e.g.][]{2012MNRAS.427..274S}.
The azimuthal component ($V_{\odot}$) is more difficult to measure, but can be
modelled using the asymmetric drift relation \citep{2008gady.book.....B}. The
values of the components of the LSR are usually taken from
\citet{2010MNRAS.403.1829S}. The value of the circular velocity is taken to be
$\sim 220\kms$ \citep[e.g.][]{2012ApJ...759..131B}.

@@

\subsection{Velocity Variations} \label{ssec:lsr_var}
We also expect that the local standard of rest (LSR) should vary as a function
of azimuth. We perform this calculation in Appendix~\ref{app:lsr} to estimate
the components of the LSR as a function of azimuth, but performing a best-fit
subtraction to correct for misalignment of the original coordinate system (as
in the previous section) is more involved. Since we find that the variation in
the LSR is less pronounced than for the midplane, and since offsets in
velocity only contribute to second order to $J_R$ and $J_z$ when a star is at
maximum amplitude in $R$ or $z$ (where the majority of the orbit is, see
Section~\ref{ssec:epi_action}), we defer this calculation to future work.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{fig/lsr.pdf}
\end{center}
\caption{The local standard of rest (LSR) as a function of azimuth at the
fiducial Solar circle ($R_{\odot} = 8.2\,\kpc$). No best-fit subtraction is
performed here as we did in the case of the midplane
(Section~\ref{ssec:local_midplane}).}
\label{fig:lsr_variations}
\end{figure*}

>>>>

In this work we discuss how the imposition of a global axisymmetric coordinate
system based on a local measurement of the midplane location and
Galactocentric distance introduces systematic error in the computation of
actions, especially for stars in the thin disk. Our main point is that the
``local'' Galactic midplane is not the same as the ``global'' Galactic
midplane. For stars relatively far from the Sun, the local midplane location
will not agree with our local midplane extrapolated onto their position, due
to relatively small-scale variations in the density of the gas and stellar
distribution. Converting positions and velocities of more distant stars from a
heliocentric to a Galactocentric coordinate system introduces a systematic
bias in the $z$ coordinate, and hence the actions computed for these stars.
The further from the solar neighborhood the target star, the more likely the
mismatch will result in large systematic uncertainty, especially in the
vertical action $J_z$. A similar
argument applies to any remaining uncertainty in measurements of the
Galactocentric radius of the Sun and to variations in the LSR.

@@

\subsection{Local Standard of Rest}
Finally, one must define the LSR, or mean velocity of
stars near the Sun relative to the Galactic center (which is defined to have
zero velocity), and the velocity of the Sun relative to the LSR. The radial
($U_{\odot}$) and vertical ($W_{\odot}$) components are computed by taking the
mean motions of different stellar groups \citep[e.g.][]{2012MNRAS.427..274S}.
The azimuthal component ($V_{\odot}$) is more difficult to measure, but can be
modelled using the asymmetric drift relation \citep{2008gady.book.....B}. The
values of the components of the LSR are usually taken from
\citet{2010MNRAS.403.1829S}. The value of the circular velocity is taken to be
$\sim 220\kms$ \citep[e.g.][]{2012ApJ...759..131B}.

@@

\subsection{Velocity Variations} \label{ssec:lsr_var}
We also expect that the LSR should vary as a function
of azimuth. We perform this calculation in Appendix~\ref{app:lsr} to estimate
the components of the LSR as a function of azimuth, but performing a best-fit
subtraction to correct for misalignment of the original coordinate system (as
in the previous section) is more involved. Since we find that the variation in
the LSR is less pronounced than for the midplane, and since offsets in
velocity only contribute to second order to $J_R$ and $J_z$ when a star is at
maximum amplitude in $R$ or $z$ (where the majority of the orbit is, see
Section~\ref{ssec:epi_action}), we defer this calculation to future work.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{fig/lsr.pdf}
\end{center}
\caption{The LSR as a function of azimuth at the
fiducial Solar circle ($R_{\odot} = 8.2\,\kpc$). No best-fit subtraction is
performed here as we did in the case of the midplane
(Section~\ref{ssec:local_midplane}).}
\label{fig:lsr_variations}
\end{figure*}
